\documentclass{article}

\begin{document}

\centerline{\sc \large A Simple Sample \LaTeX\ File}
\vspace{.5pc}
\centerline{\sc Stupid Stuff I Wish Someone Had Told Me Four Years Ago}
\centerline{\it (Read the .tex file along with this or it won't
            make much sense)}
\vspace{2pc}

The first thing to realize about \LaTeX\ is that it is not ``WYSIWYG''.
In other words, it isn't a word processor; what you type into your
.tex file is not what you'll see in your .dvi file.  For example,
\LaTeX\ will      completely     ignore               extra
spaces    within                             a line of your .tex file.

Like this.  But any carriage-returns after the first two will be
completely ignored; in other words, you
can't
add
more
space
between

\noindent Actually there is one more way, used above; for example,
{\sc this way}.  The way that you get in and out of environment varies
depending on which kind of environment you want; for example, you use
\verb|\underline| ``outside'', but \verb|\it| ``inside'';
notice \underline{this} versus {\it this}.

The real power of \LaTeX\ (for us) is in the math environment. You
push and pop out of the math environment by typing \verb|$|. For
example, $2x^3 - 1 = 5$ is typed between dollar signs as
\verb|$2x^3 - 1 = 5$|. Perhaps a more interesting example is
$\lim_{N \to \infty} \sum_{k=1}^N f(t_k) \Delta t$.

You can get a fancier, display-style math
environment by enclosing your equation with double dollar signs.
This will center your equation, and display sub- and super-scripts in
a more readable fashion:

$$\lim_{N \to \infty} \sum_{k=1}^N f(t_k) \Delta t.$$

If you don't want your equation to be centered, but you want the nice
indicies and all that, you can use \verb|\displaystyle| and get your
formula ``in-line''; using our example this is
$\displaystyle \lim_{N \to \infty} \sum_{k=1}^N f(t_k) \Delta t.$  Of
course this can screw up your line spacing a little bit.

There are many more things to know about \LaTeX\ and we can't
possibly talk about them all here.
You can use \LaTeX\ to get tables, commutative diagrams, figures,
aligned equations, cross-references, labels, matrices, and all manner
of strange things into your documents.  You can control margins,
spacing, alignment, {\it et cetera} to higher degrees of accuracy than
the human eye can percieve.  You can waste entire days typesetting
documents to be ``just so''.  In short, \LaTeX\ rules.

The best way to learn \LaTeX\ is by example. Get yourself a bunch
of .tex files, see what kind of output they produce, and figure out how
to modify them to do what you want.  There are many template and
sample files on the department \LaTeX\ page and in real life in the
big binder that should be in the computer lab somewhere.  Good luck!

\end{document}
